---
title: "a3_task3_madeline_mcewen"
author: "Madeline"
date: "10/30/2020"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(ggplot2)
```

## Read in data

```{r}

#put data into lower snake case 
urchin_data<-read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Comparing distribution of sizes between purple and red urchins

```{r}
# faceted histogram
ggplot(data=urchin_data, aes(x=size)) +
  geom_histogram(bins=15) +
  facet_wrap(~common_name)+
  labs(x="Size (cm)", y="Amount of Urchins")
```

```{r}
#boxplot
ggplot(data=urchin_data, aes(x=common_name, y=size))+
  geom_boxplot()+
  labs(x="Common Name", y="Size (cm)")
```

## Summary table with mean, median, sd, and n
- Use group_by() and summarize()

```{r}
urchin_stats <- urchin_data %>% 
  group_by(common_name) %>% 
  summarize(mean = mean(size), 
            standard_deviation = sd(size), 
            median= median(size), 
            n=n())
urchin_stats
```

Looking at the histogram ggplots, the red urchin is much more evenly spread than the purple. Red also has a smaller standard deviation which indicates the values are closer to the mean, which we can see when we look at the results from our graphs and summary table. 
While the means between species differ by only about 2.3, the range of sizes is only between 2.5 and 15. Keeping this in mind, a difference of 2.3 is actually quite large. 


## Exploring purple urchins 
- Use mutate() and case_when()

```{r}
purple_urchins<-urchin_data %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(Date=ymd(date)) %>% 
  mutate(year= year(Date)) %>% 
  mutate(site_full=case_when(
    site == "NAPL" ~"Naples Reef",
    site == "AQUE" ~"Arroyo Quemado Reef",
    site == "CARP" ~"Carpinteria Reef",
    site == "MOHK" ~"Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))
```

## Summary table 
### Contains total number of purple urchins observed by site and year
- use count()

```{r}
purple_data<-purple_urchins %>% 
  count(site, year)

purple_data
```

```{r}
ggplot(data=purple_data, aes(x=year, y=n))+
  geom_boxplot(aes(fill=site))+
                 scale_fill_manual(values= c("#C4961A","#D16103","#C3D7A4","#4E84C4","#F4EDCA"))+
  theme_bw()+
  labs(x="Year", y="Amount of purple urchins")

```





